Pre-/Post-Trikafta in CFRD

### 2024-09-16 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.qmd
Hba1c longitudinal

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /home/tim/Documents/Miscellaneous/zotero.bib
csl: /home/tim/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpmisc)
library(lubridate)
library(gtsummary)
library(rspiro)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Colleen Wood"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
load("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
# Ethnicity
chco_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
chco_demo$Hispanic <- NA
chco_demo$Hispanic[chco_demo$Hispanic.Latinx] <- "Yes"
chco_demo$Hispanic[!chco_demo$Hispanic.Latinx] <- "No"
chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]

montana_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]

uw_demo <- read.csv("/home/tim/OneDrive/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin
uw_demo <- uw_demo[, c("CFF.ID", "Hispanic")]

eth <- do.call(rbind, list(chco_demo, montana_demo, uw_demo))
colnames(eth) <- c("CFF ID", "Hispanic")

df <- left_join(df, eth)
```

@@

>>>>

---
title: "Pre-/Post-Trikafta in CFRD"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Dropbox/Miscellaneous/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
editor: source
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpmisc)
library(lubridate)
library(gtsummary)
library(knitr)
library(rspiro)
library(lmerTest)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Colleen Wood"
)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| include: false
load("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/analysis_dataset.rdata")
load("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/hba1c_dataset.rdata")
# Ethnicity
chco_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Raw/Data From CHCO/Registry/Copy of Earliest Trikafta Dates in Registry_updated 2.2.21 cc modified_cfrdyn.csv")
chco_demo$Hispanic <- NA
chco_demo$Hispanic[chco_demo$Hispanic.Latinx] <- "Yes"
chco_demo$Hispanic[!chco_demo$Hispanic.Latinx] <- "No"
chco_demo <- chco_demo[, c("CFF.ID", "Hispanic")]
montana_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/montana_demo.csv")
montana_demo$Hispanic <- montana_demo$Is.Patient.of.Hispanic.Origin
montana_demo <- montana_demo[, c("CFF.ID", "Hispanic")]
uw_demo <- read.csv("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Colleen Wood/Prepost Trikafta/Data_Cleaned/uw_demo.csv")
uw_demo$Hispanic <- uw_demo$Is.Patient.of.Hispanic.Origin
uw_demo <- uw_demo[, c("CFF.ID", "Hispanic")]
eth <- do.call(rbind, list(chco_demo, montana_demo, uw_demo))
colnames(eth) <- c("CFF ID", "Hispanic")
df <- left_join(df, eth)
```

@@

# HbA1c longitudinal analysis

```{r}
# Simple model with random intercept (can't fit random slope)
hba1cs$Change <- (hba1cs$Time > 0) * hba1cs$Time
mod <- lmer(HbA1c ~ Time + Change + (1 | `CFF ID`), data = hba1cs)
hba1c_df <- mod@frame
hba1c_df$pop_pred <- predict(mod, re.form = NA)
hba1c_df$pred <- predict(mod)
# Plot
ggplot(hba1c_df, aes(x = Time)) +
  geom_point(
    aes(y = HbA1c, colour = factor(`CFF ID`), group = factor(`CFF ID`)),
    alpha = 0.1
  ) +
  geom_line(
    aes(y = pred, colour = factor(`CFF ID`), group = factor(`CFF ID`)),
    alpha = 0.1
  ) +
  geom_line(aes(y = pop_pred)) +
  theme_bw() +
  theme(legend.position = "none")
# Print table
kable(summary(mod)$coefficients)
```
